---
title: "Final Project Draft"
output: html_document
authors: "Aimi,  Sonya, Ethan, Luca"
---

## Introduction

“The Billboard Hot 100 is the music industry standard record chart in the United States for songs, published weekly by Billboard magazine. Chart rankings are based on sales (physical and digital), radio play, and online streaming in the United States.”

The data was found on TidyTuesday and is from Data.World with the original data points found on Billboard.com and Spotify. The cases are songs from the certain week(s) in which they appeared on the Billboard 100 chart. It includes every weekly Hot 100 singles chart from Billboard.com. Relevant variables from Billboard include song, performer, instance (ordinal for which appearance on the chart it is for the song), previous_week_position, and more. Relevant variables from Spotify include spotify_genre, spotify_track_duration, danceability (double 0-1; factoring in tempo, rhythm stability, beat strength), energy (double 0-1; perceptual measure of intensity and activity), key, acousticness (double 0-1), and valence (double 0-1; “musical positiveness”), and more. Both include a song_id that we anticipate will be used for joining.


## Methodology

```{r load-packages}
library(tidyverse) 
```

```{r load-in-data}
billboard <- 
  readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/billboard.csv')
audio_features <-
  readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/audio_features.csv')
```

We are primarily interested in the factors that 
## Results


## Discussion
